Partitioned Local Depth (PaLD) Community Analyses in R

### 2025-02-03 2025-r-journal/dagostino-mcgowan.Rmd
add recent revision

In addition, the package provides a number of pertinent example data sets which may be used to demonstrate community analysis, including a synthetic data set of two-dimensional points created by \citet{gionis1clustering} to demonstrate aggregation, clustering data generated from the scikit-learn Python package \citep{pedregosa2011scikit}, data describing cognate relationships between words across 87 Indo-European languages \citep{dyen92}, data compiled by \cite{tissue} of tissue gene expressions, data emplying information from the World Values Survey \citep{inglehart2014world} on cultural values regarding family, religion, education, and institutions for several regions \citep{muthukrishna2020beyond}, and three example data sets generated for the \citet{berenhaut2022social} paper.

@@

```{r}
d <- dist(df)
cohesion_matrix(d)
```

@@

```{r}
tissue_graphs <- community_graphs(tissue_cohesion)
tissue_graph_strong <- tissue_graphs[["G_strong"]]
E <- igraph::get.edgelist(tissue_graph_strong)
tissue_strong_ties <- data.frame(
  strong_ties = apply(E, 1, paste, collapse = ",")
)
tissue_strong_ties |>
  dplyr::count(strong_ties) |>
  dplyr::arrange(-n)
```

>>>>

In addition, the package provides a number of pertinent example data sets which may be used to demonstrate community analysis, including a synthetic data set of two-dimensional points created by \citet{gionis1clustering} to demonstrate aggregation, clustering data generated from the scikit-learn Python package \citep{pedregosa2011scikit}, data describing cognate relationships between words across 87 Indo-European languages \citep{dyen92}, data compiled by \cite{tissue} of tissue gene expressions, data employing information from the World Values Survey \citep{inglehart2014world} on cultural values regarding family, religion, education, and institutions for several regions \citep{muthukrishna2020beyond}, and three example data sets generated for the \citet{berenhaut2022social} paper.

@@

```{r}
cohesion_matrix(d)
```

@@

```{r, echo=FALSE}
options(width = 20)
```

```{r}
tissue_graphs <- community_graphs(tissue_cohesion)
tissue_graph_strong <- tissue_graphs[["G_strong"]]
E <- igraph::get.edgelist(tissue_graph_strong)
tissue_strong_ties <- data.frame(
  strong_ties = apply(E, 1, paste, collapse = ",")
)
tissue_strong_ties |>
  table() |>
  sort(decreasing = TRUE)
```

### 2025-02-03 2025-r-journal/dagostino-mcgowan.tex
add recent revision

\begin{Schunk}
\begin{Sinput}
d <- dist(df)
cohesion_matrix(d)
\end{Sinput}
\begin{Soutput}
#>            A          B          C         D          E         F
#> A 0.25000000 0.18333333 0.06666667 0.0000000 0.18333333 0.0000000
#> B 0.14000000 0.24000000 0.05000000 0.0000000 0.10666667 0.0000000
#> C 0.07333333 0.07333333 0.20333333 0.0000000 0.03333333 0.0800000
#> D 0.00000000 0.00000000 0.00000000 0.2333333 0.00000000 0.1333333
#> E 0.14000000 0.10666667 0.03333333 0.0000000 0.24000000 0.0000000
#> F 0.00000000 0.00000000 0.05000000 0.1400000 0.00000000 0.2400000
#> attr(,"class")
#> [1] "cohesion_matrix" "matrix"          "array"
\end{Soutput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
graphs <- community_graphs(cohesion)
graphs[["G_strong"]]
\end{Sinput}
\begin{Soutput}
#> IGRAPH 5de0378 UNW- 6 3 --
#> + attr: name (v/c), weight (e/n)
#> + edges from 5de0378 (vertex names):
#> [1] A--B A--E D--F
\end{Soutput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
library(ggplot2)
ggplot(df, aes(x1, x2)) +
  geom_text(label = rownames(df)) +
  coord_fixed(ratio = 1) +
  xlim(c(4, 16)) +
  ylim(c(4, 16))
\end{Sinput}
\begin{figure}[H]
\begin{center}
\includegraphics[width=5in]{dagostino-mcgowan_files/figure-latex/fig1-1} \caption[Visualization of the points from the data frame `df`]{Visualization of the points from the data frame `df`}\label{fig:fig1}
\end{center}
\end{figure}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
plot_community_graphs(cohesion,
                      vertex.label.cex = 2,
                      vertex.label.dist = 0.9)
\end{Sinput}
\begin{figure}[H]
\includegraphics[width=5in,trim=0in 1in 0in 0.7in,clip]{dagostino-mcgowan_files/figure-latex/fig2-1} \caption[PaLD graph displaying the relationship between the points in data frame `df`]{PaLD graph displaying the relationship between the points in data frame `df`}\label{fig:fig2}
\end{figure}
\end{Schunk}

@@

plot_community_graphs(cohesion,
                      layout = as.matrix(df),
                      vertex.size = 100,
                      vertex.label.color = "white",
                      axes = TRUE,
                      rescale = FALSE,
                      asp = 1,
                      xlim = c(4, 16),
                      ylim = c(4, 16))
\end{Sinput}
\begin{figure}[H]
\includegraphics[width=5in,trim=0in 0.6in 0in 0.8in,clip]{dagostino-mcgowan_files/figure-latex/fig3-1} \caption[PaLD graph displaying the relationship between the points in data frame `df`, matching the original layout in Figure 1]{PaLD graph displaying the relationship between the points in data frame `df`, matching the original layout in Figure 1}\label{fig:fig3}
\end{figure}
\end{Schunk}

@@

\begin{Schunk}
\begin{figure}[H]
\includegraphics[width=5.5in,trim=0in 5in 0in 3in,clip]{fig5} \caption[Community cluster network for the tissue data]{Community cluster network for the tissue data. The line colors indicate the PaLD communities; the point colors indicate the tissue classification.}\label{fig:fig4}
\end{figure}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
tissue_graphs <- community_graphs(tissue_cohesion)
tissue_graph_strong <- tissue_graphs[["G_strong"]]
E <- igraph::get.edgelist(tissue_graph_strong)
tissue_strong_ties <- data.frame(
  strong_ties = apply(E, 1, paste, collapse = ",")
)
tissue_strong_ties |>
  dplyr::count(strong_ties) |>
  dplyr::arrange(-n)
\end{Sinput}
\begin{Soutput}
#>               strong_ties   n
#> 1           kidney,kidney 273
#> 2             colon,colon 241
#> 3 hippocampus,hippocampus 191
#> 4   cerebellum,cerebellum 186
#> 5             liver,liver  85
#> 6 endometrium,endometrium  81
#> 7       placenta,placenta   4
#> 8      kidney,endometrium   3
\end{Soutput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
french_neighbors <- igraph::neighbors(cognate_graph_strong, "French")
french_neighbors
\end{Sinput}
\begin{Soutput}
#> + 8/87 vertices, named, from b09505b:
#> [1] Italian         Ladin           Provencal       Walloon
#> [5] French_Creole_C French_Creole_D Spanish         Catalan
\end{Soutput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
cognate_cohesion["French", french_neighbors] |>
  sort(decreasing = TRUE)
\end{Sinput}
\begin{Soutput}
#>         Walloon       Provencal French_Creole_C French_Creole_D           Ladin
#>      0.03258771      0.02871174      0.02406057      0.02406057      0.02094596
#>         Italian         Catalan         Spanish
#>      0.01997696      0.01859688      0.01679733
\end{Soutput}
\end{Schunk}

@@

\begin{Schunk}
\begin{figure}[H]
\includegraphics[width=5.5in,trim=0in 5in 0in 1in,clip]{fig6} \caption[Community structure for 87 Indo-European languages, which employs cognate information that was coded via 2,665-dimensional binary vectors]{Community structure for 87 Indo-European languages, which employs cognate information that was coded via 2,665-dimensional binary vectors. Commonly identifiable language clusters arise along with informative inter- and intra-cluster structure. Several ancient languages are centrally located.}\label{fig:figlang}
\end{figure}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
par(mfrow = c(1, 3), pty = "s")
plot(
  exdata3,
  pch = 16,
  col = pald_colors[exdata_pald],
  xlab = "",
  ylab = "",
  main = "PaLD Communities",
  asp = 1
)
plot(
  exdata3,
  pch = 16,
  col = pald_colors[exdata_kmeans],
  xlab = "",
  ylab = "",
  main = "K-Means Clusters (k = 8)",
  asp = 1
)
plot(
  exdata3,
  pch = 16,
  col = pald_colors[exdata_hclust],
  xlab = "",
  ylab = "",
  main = "Hiearchical Clusters (k = 8)",
  asp = 1
)
\end{Sinput}
\begin{figure}[H]
\includegraphics[width=5.5in,trim=0in 1in 0in 0.8in,clip]{dagostino-mcgowan_files/figure-latex/fig5-1} \caption[PaLD, k-means, and hierarchical 8-clustering of randomly generated example data (from \citet{berenhaut2022social}]{PaLD, k-means, and hierarchical 8-clustering of randomly generated example data (from \citet{berenhaut2022social}; Figure 4D).}\label{fig:fig5}
\end{figure}
\end{Schunk}

@@

\begin{Schunk}
\begin{figure}[H]
\includegraphics[width=5.5in,trim=0in 5in 0in 3in,clip]{fig7} \caption[Community structure for cultural distance data]{Community structure for cultural distance data.}\label{fig:figculture}
\end{figure}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
dist_cohesion_plot(cultures, mutual = TRUE)
\end{Sinput}
\begin{figure}[H]
\includegraphics[width=5.5in,trim=0in 0.5in 0in 0.5in,clip]{dagostino-mcgowan_files/figure-latex/figco-1} \caption{A plot of cohesion versus distance for the data. The identified communities are colored as in Figure \ref{fig:figculture}.}\label{fig:figco}
\end{figure}
\end{Schunk}

>>>>

\begin{Schunk}
\begin{Sinput}
cohesion_matrix(d)
\end{Sinput}
\begin{Soutput}
#>            A          B          C         D          E         F
#> A 0.25000000 0.18333333 0.06666667 0.0000000 0.18333333 0.0000000
#> B 0.14000000 0.24000000 0.05000000 0.0000000 0.10666667 0.0000000
#> C 0.07333333 0.07333333 0.20333333 0.0000000 0.03333333 0.0800000
#> D 0.00000000 0.00000000 0.00000000 0.2333333 0.00000000 0.1333333
#> E 0.14000000 0.10666667 0.03333333 0.0000000 0.24000000 0.0000000
#> F 0.00000000 0.00000000 0.05000000 0.1400000 0.00000000 0.2400000
#> attr(,"class")
#> [1] "cohesion_matrix" "matrix"          "array"
\end{Soutput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
graphs <- community_graphs(cohesion)
graphs[["G_strong"]]
\end{Sinput}
\begin{Soutput}
#> IGRAPH cc642b5 UNW- 6 3 -- 
#> + attr: name (v/c), weight (e/n)
#> + edges from cc642b5 (vertex names):
#> [1] A--B A--E D--F
\end{Soutput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
library(ggplot2)
ggplot(df, aes(x1, x2)) +
  geom_text(label = rownames(df)) + 
  coord_fixed(ratio = 1) + 
  xlim(c(4, 16)) + 
  ylim(c(4, 16))
\end{Sinput}
\begin{figure}
\includegraphics{dagostino-mcgowan_files/figure-latex/fig1-1} \caption[Visualization of the points from the data frame `df`]{Visualization of the points from the data frame `df`}\label{fig:fig1}
\end{figure}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
plot_community_graphs(cohesion, 
                      vertex.label.cex = 2,
                      vertex.label.dist = 0.9)
\end{Sinput}
\begin{figure}
\includegraphics{dagostino-mcgowan_files/figure-latex/fig2-1} \caption[PaLD graph displaying the relationship between the points in data frame `df`]{PaLD graph displaying the relationship between the points in data frame `df`}\label{fig:fig2}
\end{figure}
\end{Schunk}

@@

plot_community_graphs(cohesion, 
                      layout = as.matrix(df),
                      vertex.size = 100,
                      vertex.label.color = "white",
                      axes = TRUE,
                      rescale = FALSE,
                      asp = 1,
                      xlim = c(4, 16),
                      ylim = c(4, 16))
\end{Sinput}
\begin{figure}
\includegraphics{dagostino-mcgowan_files/figure-latex/fig3-1} \caption[PaLD graph displaying the relationship between the points in data frame `df`, matching the original layout in Figure 1]{PaLD graph displaying the relationship between the points in data frame `df`, matching the original layout in Figure 1}\label{fig:fig3}
\end{figure}
\end{Schunk}

@@

\begin{Schunk}
\begin{figure}
\includegraphics[width=1\linewidth]{fig5} \caption[Community cluster network for the tissue data]{Community cluster network for the tissue data. The line colors indicate the PaLD communities; the point colors indicate the tissue classification.}\label{fig:fig4}
\end{figure}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
tissue_graphs <- community_graphs(tissue_cohesion)
tissue_graph_strong <- tissue_graphs[["G_strong"]]
E <- igraph::get.edgelist(tissue_graph_strong)
tissue_strong_ties <- data.frame(
  strong_ties = apply(E, 1, paste, collapse = ",")
)
tissue_strong_ties |>
  table() |>
  sort(decreasing = TRUE)
\end{Sinput}
\begin{Soutput}
#> strong_ties
#>           kidney,kidney 
#>                     273 
#>             colon,colon 
#>                     241 
#> hippocampus,hippocampus 
#>                     191 
#>   cerebellum,cerebellum 
#>                     186 
#>             liver,liver 
#>                      85 
#> endometrium,endometrium 
#>                      81 
#>       placenta,placenta 
#>                       4 
#>      kidney,endometrium 
#>                       3
\end{Soutput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
french_neighbors <- igraph::neighbors(cognate_graph_strong, "French")
french_neighbors
\end{Sinput}
\begin{Soutput}
#> + 8/87 vertices, named, from 4454cb9:
#> [1] Italian        
#> [2] Ladin          
#> [3] Provencal      
#> [4] Walloon        
#> [5] French_Creole_C
#> [6] French_Creole_D
#> [7] Spanish        
#> [8] Catalan
\end{Soutput}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
cognate_cohesion["French", french_neighbors] |>
  sort(decreasing = TRUE)
\end{Sinput}
\begin{Soutput}
#>         Walloon 
#>      0.03258771 
#>       Provencal 
#>      0.02871174 
#> French_Creole_C 
#>      0.02406057 
#> French_Creole_D 
#>      0.02406057 
#>           Ladin 
#>      0.02094596 
#>         Italian 
#>      0.01997696 
#>         Catalan 
#>      0.01859688 
#>         Spanish 
#>      0.01679733
\end{Soutput}
\end{Schunk}

@@

\begin{Schunk}
\begin{figure}
\includegraphics[width=1\linewidth]{fig6} \caption[Community structure for 87 Indo-European languages, which employs cognate information that was coded via 2,665-dimensional binary vectors]{Community structure for 87 Indo-European languages, which employs cognate information that was coded via 2,665-dimensional binary vectors. Commonly identifiable language clusters arise along with informative inter- and intra-cluster structure. Several ancient languages are centrally located.}\label{fig:figlang}
\end{figure}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
par(mfrow = c(1, 3), pty = "s")
plot(
  exdata3,
  pch = 16,
  col = pald_colors[exdata_pald],
  xlab = "",
  ylab = "",
  main = "PaLD Communities",
  asp = 1
)
plot(
  exdata3,
  pch = 16,
  col = pald_colors[exdata_kmeans],
  xlab = "",
  ylab = "",
  main = "K-Means Clusters (k = 8)",
  asp = 1
)
plot(
  exdata3,
  pch = 16,
  col = pald_colors[exdata_hclust],
  xlab = "",
  ylab = "",
  main = "Hiearchical Clusters (k = 8)",
  asp = 1
)
\end{Sinput}
\begin{figure}
\includegraphics{dagostino-mcgowan_files/figure-latex/fig5-1} \caption[PaLD, k-means, and hierarchical 8-clustering of randomly generated example data (from \citet{berenhaut2022social}]{PaLD, k-means, and hierarchical 8-clustering of randomly generated example data (from \citet{berenhaut2022social}; Figure 4D).}\label{fig:fig5}
\end{figure}
\end{Schunk}

@@

\begin{Schunk}
\begin{figure}
\includegraphics[width=1\linewidth]{fig7} \caption[Community structure for cultural distance data]{Community structure for cultural distance data.}\label{fig:figculture}
\end{figure}
\end{Schunk}

@@

\begin{Schunk}
\begin{Sinput}
dist_cohesion_plot(cultures, mutual = TRUE)
\end{Sinput}
\begin{figure}
\includegraphics{dagostino-mcgowan_files/figure-latex/figco-1} \caption{A plot of cohesion versus distance for the data. The identified communities are colored as in Figure \ref{fig:figculture}.}\label{fig:figco}
\end{figure}
\end{Schunk}
